Dark Control:\\ Towards a Unified Account of Default Mode Function\\ by Markov Decision Processes

### 2017-12-06 paper/paper.tex
reread part 1

\begin{abstract}
The Default Mode Network (DMN) is believed to subserve the
baseline mental activity in humans.
%
Its highest energy consumption compared to other brain networks and
its intimate coupling with conscious awareness are both pointing to
an overarching function.
%
Many research streams support an evolutionarily adaptive role in
envisioning experience to anticipate the future.
In the present work, we propose a \textit{process model}
that tries to explain \textit{how}
the DMN may implement
continuous evaluation and prediction of the environment to guide behavior.
%
% DMN function is recast in terms of
% control theory and reinforcement learning by basing it on Markov decision processes.
%
Specifically, we answer the question whether the neurobiological properties of the default mode network collectively provide the computational building blocks necessary for a Markov Decision Process.
%
We argue that our formal account of DMN function naturally accommodates as special cases
the previously proposed cognitive accounts on
(1) predictive coding,
(2) semantic associations, and
(3) a ``sentinel'' role.
%
Moreover, this process model for the neural optimization of complex behavior in the DMN
offers parsimonious explanations for
recent experimental findings in animals and humans.
\end{abstract}

@@

This \textit{dark matter of brain physiology}
\citep{raichle2006brain} begs the question
of the biological purpose underlying DMN activity.
Despite observation of similar large-scale networks of co-varying spontaneous activity in MEG investigations \citep{de2010temporal,brookes2011investigating,baker2014fast}, the link between the fMRI BOLD signal and population-level neural activity is still unclear. If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains challenging \citep{maldjian2014graph}.

@@

\subsection{Towards a formal account of default mode function: higher-order control of the organism}
The network nodes that compose the human DMN
are responsible for extended parts of the baseline neural activity,
which typically decreases when engaged in psychological experiments
\citep{raichle_baseline}.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
Today, many psychologists and neuroscientists believe that
the DMN implements some form of
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008, spreng2009common}.
This brain network
might have emerged to continuously predict the environment using
mental imagery as an evolutionary advantage \citep{suddendorf2007evolution}.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode regions \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into DMN function is
particularly challenging because
this brain network appears to
simultaneously modulate perception-action cycles in the present
and to support mental travel
across time, space, and content domains
\citep{boyer2008evolutionary}.

@@

Psychologically,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
Conversely,
it has been empirically shown that, the more the world is easy to foresee,
the more human mental activity
becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation, mason2007, weiss2006}.
Without requiring explicit awareness,
these ``offline'' processes may contribute
to optimizing control of the organism.
We formalize
a \textit{policy matrix} to capture the space of possible actions that
the agent can perform
on the environment given the current state. A \textit{value function}
maps environmental objects and events (i.e., states) to expected rewards.
Switching between states reduces to a sequential processing model.
Informed by outcomes of performed actions, neural computation reflected in
DMN dynamics could be constantly shaped by prediction error through feedback loops.
This computational account of DMN will be described in the mathematical framework of Markov Decision Processes (MDPs). MDPs indeed address specifically the case of decision making in stochastic context with rewards.

Such an RL account of DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.

We argue that DMN implication in many
advanced cognitive processes in humans
can be recast as prediction error minimization
based on probabilistic mental simulations
- ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
thus maximizing action outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide blocks of pseudo-experience to iteratively
optimize the behavior of the organism.

Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
the human brain is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
The present paper proposes a
process model that satisfies this contention.
We also contribute to the discussion of DMN function
by providing some of the first empirical evidence that
morphological variability in DMN regions
is linked to the reward circuitry (Fig. 2).
Finally, we detail how our process model relates
to previous accounts of DMN function and
we derive explicit hypotheses to be tested in future neuroscience
experiments (Box 2).
At this stage, we emphasize the importance of differentiating which levels of observation are at play in the present account. A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent. Henceforth, we will use ``inference`` when describing aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.

\section{Known neurobiological properties of the default mode network}
\label{sec:nodes}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in the neuroscience literature.
This walkthrough across each region of the DMN will
outline the individual functional profiles,
paving the way for their algorithmic interpretation
in our formal account (section 3).

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{\textbf{Default mode network: key functions.}
  Neurobiological overview of the DMN with its major constituent nodes and
  the associated functional roles relevant in our functional interpretation.}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the human DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are probably responsible for the highest turn-over
of energy consumption \citep{raichle2001pnas, raichle_baseline}.
These metabolic characteristics go hand-in-hand with
neuroimaging analyses that suggested the PMC and mPFC to
potentially represent the functional backbone of the DMN
\citep{andrews2010, hagmann2008mapping}.

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication.
%
This DMN region has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from medial damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Neurological patients are plagued by an
inability to combine various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological relevance assessment
\citep{amygdala_bzdok}.

The putative role of the PMC in continuous abstract integration of
environmental relevance
and ensuing top-level guidance of action on the environment is supported
by many neuroscience experiments.
Electrophysiological recordings in animals implicated PMC neurons in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
outcome-dependent behavioral modulation \citep{hayden2009electrophysiological},
as well as approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would pursue an exploratory or exploitative
behavioral strategy during food foraging \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN region encoded the
preference for or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its actual value,
based on an ``internal currency of value'' \citep{mccoy2005risk}.
In fact, direct stimulation of PMC neurons
promoted exploratory actions,
which would otherwise be shunned \citep{hayden2008stim}.
Graded changes in firing rates of PMC neurons
indicated changes in upcoming choice trials, while their neural patterns were
distinct from neuronal spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different parts of auditory narratives in an fMRI study
\citep{simony2016dynamic}.

Moreover, the retrosplenial portion of the PMC could support
representation of action possibilities
and evaluation of reward outcomes by integrating
information from memory recall and different perspective frames.
Regarding memory recall, retrosplenial damage has been
consistently associated with anterograde and retrograde memory impairments
of various kinds of sensory information
in animals and humans
\citep{vann2009does}.
Regarding perspective frames, the retrosplenial subregion of the PMC has been
proposed to mediate between the organism's egocentric
(i.e., focused on sensory environment) and
allocentric (i.e., focused on world knowledge) viewpoints
in animals and humans
\citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible actions and integrating that information
with memory and perspective frames
into short- and longer-term behavioral agendas.
Estimated value, that differs across individuals, might enrich
statistical assessment of the environment
to map and predict delayed reward opportunities in the future.
In doing so, the PMC may continuously adapt the organism to changes
in both the external environment and its internal representation
to enable strategic behavior.

\subsection{The prefrontal cortex: action consideration and stimulus-value association}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
multi-sensory processes
across time, space, and content domains to
exert top-level control on behavior.
Comparing to the PMC, however,
dmPFC function may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as it
potentially subserves the de-novo construction and manipulation
of meaning representations instructed by stored semantics and memories
\citep{bzdok2013segregation}.
The dmPFC may subserve inference, representation, and assessment
of one's own and other individuals' action considerations.
Generally,
neurological patients with tissue damage in the prefrontal cortex
are known to struggle with
adaptation to new stimuli and events
\citep{stuss1986frontal}.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
outcomes of these predictions.
Neural activity in the dmPFC indeed explained the performance decline
of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
Certain dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
% In fact, the spatially neighboring dorsal anterior cingulate cortex
% has also been linked to computing values and efforts of
% persisting a behavioral plan versus switching the
% environmental context in several lesion studies \citep{kolling2016value}.
%
% Such highly abstract neural computations necessarily rely on the
% construction of probabilistic internal information drawing from
% memory recall, mental scene elaboration processes,
% and stored knowledge of world regularities.
% %
% Moreover,
% in a computational neuroimaging experiment,
% dorsomedial PFC activity preferentially processed the consequences of
% action choices that were considered but not actually executed,
% whereas ventromedial PFC (vmPFC) activity
% processed especially value outcomes of actions actually performed on the environment
% \citep{nicolle2012agent}.

Comparing to the dmPFC,
the vmPFC probably subserves
subjective value evaluation and
risk estimation of relevant environmental stimuli
(see Fig. \ref{fig:neurobiological_overview_DMN} and \ref{fig:dmn}).
The ventromedial prefrontal DMN may
subserve adaptive behavior by bottom-up-driven
processing of “what matters now”,
drawing on sophisticated value representations
\citep{rolls_OFC, doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based action choice
\citep{glascher2012lesion}.
Indeed,
this DMN region is preferentially connected with reward-related and limbic regions.
The vmPFC is well known to have direct connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals
from dopaminergic neurotransmitter pathways
\citep{schultz1998predictive}
that not only channel action towards basic survival needs but also
enable more abstract reward processings, and thus perhaps RL,
in humans \citep{doherty2015structure}.

% \begin{wrapfigure}{r}{.4\linewidth}
\begin{figure}[!htbp]
  \begin{minipage}[c]{0.63\textwidth}
    \includegraphics[width=.8\linewidth]{fig_dmn.png}
  \end{minipage}
  \hspace{-4em}
  \begin{minipage}[c]{0.45\textwidth}
    \caption{\textbf{Morphological coupling between reward system and
        default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from volume in the DMN regions. These out-of-sample generalization performances were obtained from support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure.
      Consistent for the left and right reward system, NAc volume in a given subject is
      positively coupled with the vmPFC and HC. The colors are indicative of the (\textcolor{red}{red} = positive,
      \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  Code and data for reproduction and visualization:
  \url{www.github.com/banilo/to_be_added_later}.
  }
  \label{fig:dmn}
\end{minipage}
% \end{wrapfigure}
\end{figure}

Consistently, diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative}
quantified the NAc to
be more connected to the vmPFC than dmPFC in both species.
Two different functional connectivity analyses in humans also strongly connected
the vmPFC with the NAc, hippocampus (HC),
and PMC \citep{bzdok2015subspecialization}.
%
In line with these connectivity findings in animals and humans,
the vmPFC is often proposed to represent triggered
emotional and motivational states \citep{damasio1996somatic}.
Such real or imagined arousal states could be mapped in the vmPFC
as a bioregulatory disposition influencing cognition
and decision making.
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
value estimates
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in the vmPFC \citep{Schuck2016}.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from
the sensory environment.

@@

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory perception.
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly advanced cognitive processes
can be parsimoniously explained as probabilistic mental
scene of experience simulations coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of MCMC sampling.

@@

\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice. Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before making analogous choices in the future. Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.

>>>>

\begin{abstract}
The Default Mode Network (DMN) is believed to subserve the
baseline mental activity in humans.
%
Its highest energy consumption compared to other brain networks and
its intimate coupling with conscious awareness are both pointing to
an overarching function.
%
Many research streams speak in favor of an evolutionarily adaptive role in
envisioning experience to anticipate the future.
In the present work, we propose a \textit{process model}
that tries to explain \textit{how}
the DMN may implement
continuous evaluation and prediction of the environment to guide behavior.
%
% DMN function is recast in terms of
% control theory and reinforcement learning by basing it on Markov decision processes.
%
Specifically, we answer the question whether the neurobiological properties of the default mode network collectively provide the computational building blocks necessary for a Markov Decision Process.
%
We argue that our formal account of DMN function naturally accommodates as special cases
the previously proposed cognitive accounts on
(1) predictive coding,
(2) semantic associations, and
(3) a ``sentinel'' role.
%
Moreover, this process model for the neural optimization of complex behavior in the DMN
offers parsimonious explanations for
recent experimental findings in animals and humans.
\end{abstract}

@@

This \textit{dark matter of brain physiology}
\citep{raichle2006brain} begs the question
of the biological purpose underlying DMN activity.
Despite observation of similar large-scale networks of co-varying spontaneous activity
in electrophysiological investigations \citep{de2010temporal,brookes2011investigating,baker2014fast}, the link between the fMRI BOLD signal and population-level neural activity is still unclear. If those frequency-specific electrophysiological correlations are proposed as complementary to those observed with BOLD \citep{hipp2015bold}, their role in DMN function remains elusive \citep{maldjian2014graph}.

@@

\subsection{Towards a formal account of default mode function: higher-order control of the organism}
The network nodes that compose the human DMN
are responsible for extended parts of the baseline neural activity,
which typically decreases when engaged in controlled psychological experiments
\citep{raichle_baseline}.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
Today, many psychologists and neuroscientists believe that
the DMN implements some form of
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008, spreng2009common}.
This brain network
might have emerged to continuously predict the environment using
mental imagery as an evolutionary advantage \citep{suddendorf2007evolution}.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode regions \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into DMN function is
particularly challenging because
this brain network appears to
simultaneously modulate perception-action cycles in the present
and to support mental travel
across time, space, and content domains
\citep{boyer2008evolutionary}.

@@

Psychologically,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
Conversely,
it has been empirically shown that, the more the world is easy to foresee,
the more human mental activity
becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation, mason2007, weiss2006}.
Without requiring explicit awareness,
these ``offline'' processes may contribute
to optimizing control of the organism.
We formalize
a \textit{policy matrix} to capture the space of possible actions that
the agent can perform
on the environment given the current state. A \textit{value function}
maps environmental objects and events (i.e., states) to expected rewards.
Switching between states reduces to a sequential processing model.
Informed by outcomes of performed actions, neural computation reflected in
DMN dynamics could be constantly shaped by prediction error through feedback loops.
Such an RL account of DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.

We argue that DMN implication in many
advanced cognitive processes in humans
can be recast as prediction error minimization
based on probabilistic mental simulations
- ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
thus maximizing action outcomes across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling.
Even (unavoidably) imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide randomly sampled blocks of pseudo-experience instrumental to iteratively
optimize the behavior of the organism.

Evidence from computational modeling of human behavior \citep{kording2004bayesian} and cell recording experiments in ferrets \citep{fiser2004small} suggest that
the brain is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
The present paper proposes a
process model that satisfies this previously proposed contention.
We also contribute to the discussion of DMN function
by providing some of the first empirical evidence that
morphological variability in DMN regions
is linked to the reward circuitry (Fig. 2),
thus linking two literatures with currently scarce cross-references.
Finally, we detail how our process model relates
to previous accounts of DMN function and
we derive explicit hypotheses to be tested in future neuroscience
experiments (Box 2).
At this stage, it is important to clarify how different levels of observation
are at play in the current account. A process model is not solely intended to 
capture behavior of the agent but also the neurocomputational specifics of the agent.
In the following, we will use the word ``inference'' when describing the formal aspects
of the statistical model, the word ``prediction'' when discussing the neural implementation,
and words like ``forecast'' or ``forsee'' when refering to the behavior of the agent.

\section{Known neurobiological properties of the default mode network}
\label{sec:nodes}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in the neuroscience literature.
This walkthrough across main regions of the DMN will
outline the individual functional profiles,
paving the way for their algorithmic interpretation
in our formal account (section 3).

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{\textbf{Default mode network: key functions.}
  Neurobiological overview of the DMN with its major constituent nodes and
  the associated functional roles relevant in our functional interpretation.}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the human DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are probably responsible for highest turn-overs
of energy consumption \citep{raichle2001pnas, raichle_baseline}.
These metabolic characteristics go hand-in-hand with
neuroimaging findings that suggested the PMC and mPFC to
potentially represent the functional core of the DMN
\citep{andrews2010, hagmann2008mapping}.

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication.
%
This DMN region has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from medial damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Such neurological patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This dysfunction can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two different functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological relevance assessment
\citep{amygdala_bzdok, ghods2009fundamental, baxter2002amygdala}.

The putative role of the PMC in continuous abstract integration of
environmental relevance
and ensuing top-level guidance of action on the environment is supported
by many neuroscience experiments.
Electrophysiological recordings in animals implicated PMC neurons in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
outcome-dependent behavioral modulation \citep{hayden2009electrophysiological},
as well as approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would pursue an exploratory or exploitative
behavioral strategy during food foraging \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN region encoded the
preference for or aversion to options with uncertain reward outcomes
and its neural spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its actual value,
based on an ``internal currency of value'' \citep{mccoy2005risk}.
In fact, direct stimulation of PMC neurons in monkeys
promoted exploratory actions,
which would otherwise be shunned \citep{hayden2008stim}.
Graded changes in firing rates of PMC neurons
indicated changes in upcoming choice trials, while their neural patterns were
distinct from neuronal spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different parts of auditory narratives in an fMRI study
\citep{simony2016dynamic}.

Moreover, the retrosplenial portion of the PMC could support
representation of action possibilities
and evaluation of reward outcomes by integrating
information from memory recall and different perspective frames.
Regarding memory recall, retrosplenial damage has been
consistently associated with anterograde and retrograde memory impairments
of various kinds of sensory information
in animals and humans
\citep{vann2009does}.
Regarding perspective frames, the retrosplenial subregion of the PMC has been
proposed to mediate between the organism's egocentric
(i.e., focused on external sensory environment) and
allocentric (i.e., focused on internal world knowledge) viewpoints
in animals and humans
\citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible actions and integrating that information
with memory and perspective frames
into short- and longer-term behavioral agendas.
Estimated value, found to differs across individuals, might enrich
statistical assessment of the environment
to map and predict delayed reward opportunities in the future.
In doing so, the PMC may continuously adapt the organism to changes
in both the external environment and its internal representation
to enable strategic behavior.

\subsection{The prefrontal cortex: action consideration and stimulus-value association}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
multi-sensory processes
across time, space, and content domains to
exert top-level control on behavior.
Comparing to the PMC, however,
dmPFC function may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as this DMN part
potentially subserves the de-novo construction and manipulation
of meaning representations instructed by stored semantics and memories
\citep{bzdok2013segregation}.
The dmPFC may subserve representation and assessment
of one's own and other individuals' action considerations.
Generally,
neurological patients with tissue damage in the prefrontal cortex
are known to struggle with
adaptation to new stimuli and events
\citep{stuss1986frontal}.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
outcomes of these predictions.
Neural activity in the dmPFC indeed explained the performance decline
of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
Certain dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, action
with fine-grained adjustment of contextual aspects \citep{yoshida2010neural}.
% In fact, the spatially neighboring dorsal anterior cingulate cortex
% has also been linked to computing values and efforts of
% persisting a behavioral plan versus switching the
% environmental context in several lesion studies \citep{kolling2016value}.
%
% Such highly abstract neural computations necessarily rely on the
% construction of probabilistic internal information drawing from
% memory recall, mental scene elaboration processes,
% and stored knowledge of world regularities.
% %
% Moreover,
% in a computational neuroimaging experiment,
% dorsomedial PFC activity preferentially processed the consequences of
% action choices that were considered but not actually executed,
% whereas ventromedial PFC (vmPFC) activity
% processed especially value outcomes of actions actually performed on the environment
% \citep{nicolle2012agent}.

Comparing to the dmPFC,
the vmPFC is probably more specifically devoted to
subjective value evaluation and
risk estimation of relevant environmental stimuli
(Fig. \ref{fig:neurobiological_overview_DMN} and \ref{fig:dmn}).
The ventromedial prefrontal DMN may
subserve adaptive behavior by bottom-up-driven
processing of “what matters now”,
drawing on sophisticated value representations
\citep{rolls_OFC, doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based action choice
\citep{glascher2012lesion}.
Indeed,
this DMN region is preferentially connected with reward-related and limbic regions.
The vmPFC is well known to have direct connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is further thought to reflect reward prediction signals
from dopaminergic neurotransmitter pathways
\citep{schultz1998predictive}
that not only channel action towards basic survival needs but also
enable more abstract reward processings, and thus perhaps RL,
in humans \citep{doherty2015structure}.

% \begin{wrapfigure}{r}{.4\linewidth}
\begin{figure}[!htbp]
  \begin{minipage}[c]{0.63\textwidth}
    \includegraphics[width=.8\linewidth]{fig_dmn.png}
  \end{minipage}
  \hspace{-4em}
  \begin{minipage}[c]{0.45\textwidth}
    \caption{\textbf{Morphological coupling between reward system and
        default mode network.} Based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from volume in the DMN regions. These out-of-sample generalization performances were obtained from support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure.
      Consistent for the left and right reward system, NAc volume in a given subject is
      positively coupled with the vmPFC and HC. The colors are indicative of the (\textcolor{red}{red} = positive,
      \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  Code for reproduction and visualization:
  \url{www.github.com/banilo/darkcontrol_PCB2018}.
  }
  \label{fig:dmn}
\end{minipage}
% \end{wrapfigure}
\end{figure}

Consistently, diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative}
quantified the NAc to
be more connected to the vmPFC than dmPFC in both species.
Two different functional connectivity analyses in humans also revealed
strong vmPFC connections
with the NAc, hippocampus (HC),
and PMC \citep{bzdok2015subspecialization}.
%
In line with these connectivity findings in animals and humans,
the vmPFC is often proposed to represent triggered
emotional and motivational states \citep{damasio1996somatic}.
Such real or imagined arousal states could be mapped in the vmPFC
as a bioregulatory disposition influencing cognition
and decision making.
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
value predictions
\citep{behrens2008associative},
which may also explain why performance within and across participants
was reported to relate to state encoding in the vmPFC \citep{Schuck2016}.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from
the sensory environment.

@@

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory perception.
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly advanced cognitive processes
can be parsimoniously explained as probabilistic mental
scene of experience simulations coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of MCMC sampling
\citep{tenenbaum2011grow}.

@@

\item Experiment (Animals): We hypothesize that experience replay for browsing problem solutions subserved by the DMN is necessary for choice behavior in mice. Hippocampal single-cell recordings have shown that neural patterns during experimental choice behavior are reiterated during sleep and before make analogous choices in the future. Necessity of the DMN, in addition to the hippocampus, for “mind-searching” actions during choice behavior can be demonstrated by causal disruption of DMN regions, such as circumscribed brain lesion or optogenetic intervention in the inferior parietal and prefrontal cortices.
